visdat: Visualising whole data frames

### 2017-07-03 paper/response.Rmd

## Common issues

> It is unintuitive to me to have the rows in reverse order (i.e. row 1 at the bottom) and for the columns to be clustered by type, rather than appear in the order they appear in the data frame. I think the default behavior should be for rows and columns to appear in the same order as in the input. Additionally, always putting the titles at the top of the columns makes sense to me. Including sort_type as an option might be useful, but by default it should be off. I'm not quite sure of the use case for the flip argument, but maybe there is one!

> The flip argument should be available for all of the functions if you provide it for any, for the sake of consistency. Possibly also the palette argument.

- I agree that the rows should be in the order that you see the dataframes. I would also prefer for the column names to be on top of the visualisation - I have now implemented this, although I think that the presentation of the column names is perhaps not as nice.

@@

```

- Thank you for this, I have created a helper function for these code duplications, it makes things much easier to handle! These are now in the [internals file](https://github.com/njtierney/visdat/blob/fix-functionality/R/internals.R)

> vis_compare and vis_guess are indicated as being in beta, which seems also to apply to the plotly versions of the functions. The message that they emit is a helpful indication that they may change in the future. Before submitting to CRAN, however, you might consider moving those functions to the development version of the package and only uploading the functions with a stable API, then adding the beta version later once they stabilize. This is a judgment call; I think I tend towards being conservative on this issue personally.

- Agreed, I will create a dev branch, where these experimental features can reside.

@@

> This is a minor issue, but was so thoroughly drummed into me by CS professors that I have to mention it. In calls like if (sort_type == TRUE), the == TRUE is redundant, so it's equivalent to just write if (sort_type). Similarly, if (x == FALSE) is equivalent to if (!x). There are a number of these in the code for various functions.

- I've actually not heard this before, but I can certainly understand why you would want to avoid called such as if(x == TRUE) over if(x). This has now been changed.

@@

# vis_dat()

> Maximum data frame size: the heatmaps stop working on my computer with very large data sets. E.g. library(nycflights13); vis_dat(flights) hangs for a minute or so and then displays an empty grid. It works with 10000 rows, but not 100000. Similar issue with many columns. Especially because this package is designed for people to get a sense of new data sets, where users may not have realized how big the data are, I think it makes sense to prevent this issue. Possibilities: (1) max_rows = 10000 parameter, which stops if nrow(data) > max_rows and gives a descriptive error message. Then the user can increase the maximum or subset their data frame. Might also need to consider max_cols because 10000 rows and 2 cols is very different from 10000 rows and 500 cols. (2) Downsample the non-missing rows somehow and indicate to the user where omissions were made. (3) ???

- This is an interesting one! It was brought to my attention [here](https://github.com/njtierney/visdat/issues/32), by @MilesMcBain, but I couldn't replicate this issue. Similarly, I can actually display the nycflights data. So it seems that this is to do with the fact that computations power has a bit to do with it. I think that option 1 is the best case here, but option 2 actually gets a little bit difficult, because this leads to another issue of effectively downsampling a dataset whilst maintaining data structure. This could be a problem, say, if I did perform down-sampling and then the user looked at the data with visdat and plotly, they might get a mis-represented sample. Likewise, if there is a very large proportion of missing data, this might be difficult to preserve exactly. Ways around this might include some very explicit warning messages, or perhaps even a note directly on the plot that the data have been downsampled. ^At the moment one solution forward is to have a check first, something like:

```r
vis_dat(really_large_data)

Warning: This dataset contains over 10,000 rows, which may take a while to render. Are you sure that you want to continue?

1: Yes
2. No (You should use a sample of your data)
```

What do you think?

@@

> The palette argument doesn't cause color changes, as far as I can tell. I think you need to do vis_dat_plot <- vis_dat_plot + <palette>, whereas currently the value of vis_dat_plot isn't being updated in the blocks related to color palettes. Additionally, you could do ggplot2::scale_color_brewer(type = "qual", palette = "Set1") rather than explicitly writing out the color names, although the colors would be in a slightly different order than you have now.

- ^I believe that this is now fixed,

```r
vis_dat(airquality, palette = "qual")
vis_dat(airquality, palette = "cb_safe")
```

Return different plots now, right?

@@

> If the user provides a palette name, you should check that it's valid. If not, throw an error with a message. Currently, if you misspell a palette name, it just goes ahead with the default palette.

- Great point! I have fixed this now with:

```r
     } else  {
       warning("palette arguments need to be either 'qual' 'cb_safe' or 'default'")
     } # close else brace

@@

- great suggestion, this has been added! And can be turned off in the options with `show_perc_col == FALSE`.

## vis_compare()

> This is a neat function! It would be cool if you could generalize it to also handle data frames of unequal dimensions, perhaps by showing which rows/columns are only in one data frame or the other. Just an idea I had though, not a requirement at all!

- I am glad you like this! It is a difficult problem that I cannot work out, re the rows/cols matching when they are out of the dimension. I'm not quite sure what to do here to make this possible, I think that this should be tackled in a future release of `visdat`, and will be moved to a development branch.

@@

> The pop-up window is great! Can you make it the same as for ggplotly(vis_dat())? It's especially helpful to show the row number.

- I can get it close! Getting these the same will

> Why not have the same arguments for this function as for vis_miss?

- At the moment this is because I do now know the API for plotly that well. Once we are happy with the default arguments and appearance for vis_miss, we can focus on getting the same appearance with plotly

@@

## vis_dat_ly()

> With vis_dat(airquality) %>% ggplotly(), the window that appears on mouseover should say "type" instead of "valueType", "variable" instead of "variables" and "row" instead of "rows"

- I can get this to say "variable" instead of "variables", but I cannot change "valueType", as I think that changing this makes breaking changes. "rows" appears to be autogenerated from the `plotly::ggplotly` code, so I cannot change this.

@@

> If you provide vis_miss_ly, it makes sense to me to also provide vis_dat_ly, and possibly also vis_compare_ly and vis_guess_ly. An initial implementation could just be vis_dat(x) %>% ggplotly(), but then you have the API in place at least.

- I have whipped up `vis_dat_ly`, `vis_compare_ly`, and `vis_guess_ly` using `ggplotly(vis_*)`.

- I would prefer to build it in straight plotly, as plotly is really fast, and the process of converting ggplot plots into plotly is a bit computationally intense. I will see what I can whip up in plotly in the future. I do not know much about it at the moment, but @carson has a free book, so I'll get onto that.

@@

- Agreed, this has now been fixed.

> When running the first example from README, (vis_dat(airquality)), the user will get a warning re. deprecation of dmap(). If this is not of consequence to the package's functionality, then you might consider adding it to the suppressed messages. If it's something most users supress locally, then you might suggest that somewhere in the README.

```
#> dmap() is deprecated. Please use the new colwise family in dplyr.
#> E.g., summarise_all(), mutate_all(), etc.
```

- Thank you for finding this, I have fixed this by suppressing the error as @seaaan suggested.

@@

> Minor suggestion (which may or may not work comply with the actual vignette formatting) the purpose of demonstrating the difference between vis_guess() and vis_dat, you might consider showing the two side by side (perhaps have the vis_guess() example alone and then side by side). For this you could just add fig.show='hold', out.width='50%' to the chunk options in the Rmd.

- Excellent suggestion, done!

> Based on the README, the status of vis_dat_ly isn't totally clear (since it's not yet a function of its own).

- I have removed the mentions of these, and put them into the experimental features vignette, which is referenced in the README.

@@

> I didn't quite understand the Visualising expectations section. That said, I've also never used expectation in assertr— so, this may fall under the category of things that those who need to know do, and, thus, no more need be said.

- I've constructed a small example to illustrate this at the end of the experimental features vignette.

@@

## Documentation: Vignettes

> Some of the same copy edits I made in README I would also suggest in the vignettes (e.g. the compare this to...Where sentence that sort of goes around the vis_dat figure in the vis_guess example section)

> The experimental features vignette is a bit hard to follow at the moment, which may be in part due to the layout (the warning messages take up a good bit of space). The figure is also quite small. I'd recommend adjusting the settings to match your main vignette to make it more legible.

- Agreed, the warning messages have been fixed, and the plot size changed as well.

@@

> Note: the function examples in R all throw warning: dmap() is deprecated.
As noted in @seaaan's comments, there's also the warning re. dataframes of different dimensions
e.g. example("vis_compare") results in

```
  #>  Warning messages:
  #> 1: In if (dim(df1) != dim(df2)) { :
  #>  the condition has length > 1 and only the first element will be used
  #> 2: attributes are not identical across measure variables; they will be dropped
  #> 3: attributes are not identical across measure variables; they will be dropped

```

- Agreed, this error message for vis_compare has been addressed.

@@

> vis_compare example also seems to cluster by default (all different data is shown at the bottom, of the two columns)

- I believe this is because I changed those particular values, I'll change the values to be something else now, so that it is clear that `vis_compare()` identifies values...

@@

> Warning messages can be scary, so it's always a relief see them acknowledged or shown in vignette and/or README examples (as is done re. non-identical attributes). vis_compare (a function for which I can imagine plenty of use cases) is currently returning enough text that it may leave the user confused as to what's most important, or an error on their end, etc. Currently, running vis_compare(iris_diff, iris) (as in the given example), returns:

```
#> vis_compare is still in BETA! If you have suggestions or errors,
#>        post an issue at https://github.com/njtierney/visdat/issues
#>  dmap() is deprecated. Please use the new colwise family in dplyr.
#>  E.g., summarise_all(), mutate_all(), etc.
#>  Warning messages:
#>  1: In if (dim(df1) != dim(df2)) { :
#>    the condition has length > 1 and only the first element will be used
#>  2: attributes are not identical across measure variables; they will be dropped
#>  3: attributes are not identical across measure variables; they will be dropped
```

> The function runs fine, and appears to have worked, but it's enough "warning" text, that it may warrant concern for the user.

- Absolutely agreed, this is a problem and barrier, the messages for using `vis_compare` have been reduced to:

```r
vis_compare(typical_data)
...
```

@@

> vis_guess is where the cell-to-row visual disparity becomes most confusing. I think this is a fantastic function, especially for beginners, and so the ability to essentially "read" the visual, as one would a table (top to bottom), would be invaluable.

- Agreed, this has been addressed by making the "flip" argument become the hard set default. I cannot honestly see a good reason to use the old print method, so I have removed it to make it read top down first.

> vis_dat with plotly / interactive functions are great, and, again, here it would be even more helpful if one could "read" the charts as one would a table.

- This has been implemented for all but `vis_miss_ly` - as this is written in plotly, and I'm not certain how to get it to flip.

@@

# njtierney summary response and remaining tasks to complete

This has been a fantastic process to be a part of. To summarise, the main changes that I made were:

- re-orienting the vis_* family to read in the way that one reads a dataframe
- removing many of the error messages that occur due to some tidy/dplyr collation of factors
- refactoring repeated code into internal functions to more clearly express what was happening, and avoiding repetition.
-

@@

Still remaining to do is:

- Make a separate branch for the experimental features.
  - vis_compare()
  - vis_guess()
  - vis_*_ly() family
- complete vis_guess() for release
    - Adding the fast Rcpp code for guess_type
- Add interactivity using pure plotly
  - create `vis_dat_ly`, `vis_miss_ly`, and `vis_guess_ly` functions in plotly
- Ensure that all of the `vis_*` family functions have the same internal arguments where appropriate:

- For the moment I have decided to put visdat onto CRAN, with the main functions
being: 
  - vis_dat
  - vis_miss
- Then, in a subsequent release, I will add the functions:
  - vis_guess (speed improvements required)
  - vis_compare
  - vis_dat_ly
  - vis_miss_ly
  - vis_guess_ly
  - vis_compare_ly

@@

> example(vis_guess) / example(vis_compare) : gives warning message

- These now gives the message:

```
vis_compare / (vis_guess) is in BETA! If you have suggestions or errors
post an issue at https://github.com/njtierney/visdat/issues
```
Do you think this is OK?

- note, find all "?" "..." and "^", and paste them down here.
